---
title: "Physical Activity and Perceived Health"
subtitle: "An exploration of self-reported health across age, sex, and exercise status"
author: "Romy Bos"
date: today
format:
  html:
    theme: cosmo
---

## Overview

Using data from the **2024 CDC Behavioral Risk Factor Surveillance System (BRFSS)**, this project investigates how participating in physical exercise outside of work relates to self-reported general health across different age groups and biological sexes.

```{r}
#| label: setup
#| echo: false
#| message: false
#| warning: false

library(tidyverse)
library(scales)
library(plotly)

# Read summarized 2024 BRFSS data
data <- readRDS("clean_brfss.rds") |>
  mutate(
    tooltip_text = paste0(
      "<b>Age Group:</b> ", age_group, "<br>",
      "<b>Sex:</b> ", sex, "<br>",
      "<b>Physical Activity:</b> ", exercise, "<br>",
      "<b>Good/Excellent Health:</b> ", percent(pct_good_health, accuracy = 0.1), "<br>",
      "<b>Sample Size:</b> ", comma(total)
    )
  )
```

## Health by Exercise Status and Age

Respondents were asked whether they participated in any physical activity or exercise in the past 30 days and how they would rate their general health (*Excellent*, *Very Good*, *Good*, *Fair*, or *Poor*). *Hover over the bars to explore specific percentages and sample counts.*

```{r}
#| label: health-plot
#| fig-width: 9
#| fig-height: 5.5
#| echo: false
#| message: false
#| warning: false

p <- ggplot(data, aes(x = age_group, y = pct_good_health, fill = exercise, text = tooltip_text)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  facet_wrap(~ sex) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 0.75)) +
  scale_fill_manual(
    values = c("Any Exercise" = "#2b8cbe", "No Exercise" = "#e34a33")
  ) +
  labs(
    title = "Self-Reported Good Health by Exercise and Age",
    x = "Age Group",
    y = "Percent in Excellent or Very Good Health",
    fill = "Physical Activity"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold", size = 11)
  )

ggplotly(p, tooltip = "text")
```

## Geographic Variation Across States

Physical activity levels vary considerably across the United States. The interactive map below shows the percentage of adults reporting no physical activity in the past 30 days, along with state obesity rates.

```{r}
#| label: state-map
#| echo: false
#| message: false
#| warning: false

state_data <- readRDS("state_brfss.rds")

# Configure USA map layout
geo_config <- list(
  scope = "usa",
  projection = list(type = "albers usa"),
  showlakes = TRUE,
  lakecolor = toRGB("white")
)

# Create interactive choropleth map
plot_geo(state_data, locationmode = "USA-states") |>
  add_trace(
    z = ~inactivity_pct,
    text = ~paste0(
      "<b>", state, "</b><br>",
      "Physical Inactivity: ", inactivity_pct, "%<br>",
      "Obesity Rate: ", obesity_pct, "%"
    ),
    hoverinfo = "text",
    locations = ~state_abbr,
    color = ~inactivity_pct,
    colors = "Blues"
  ) |>
  colorbar(title = "Inactivity %") |>
  layout(
    title = list(text = "Physical Inactivity Rate by U.S. State (2023 BRFSS)"),
    geo = geo_config
  )
```

## Inactivity and Obesity Across States

Does higher physical inactivity relate to higher obesity at the state level? The scatter plot below compares the two variables across all 50 states and D.C., with a linear trend line.

```{r}
#| label: scatter-plot
#| echo: false
#| message: false
#| warning: false

p_scatter <- ggplot(
  state_data,
  aes(
    x = inactivity_pct,
    y = obesity_pct,
    text = paste0("<b>", state, "</b><br>Inactivity: ", inactivity_pct, "%<br>Obesity: ", obesity_pct, "%")
  )
) +
  geom_point(color = "#2b8cbe", size = 2.5, alpha = 0.8) +
  geom_smooth(method = "lm", se = FALSE, color = "#e34a33", linetype = "dashed", linewidth = 0.8) +
  labs(
    title = "Physical Inactivity vs. Obesity Rate by State",
    x = "Adults Inactive in Past 30 Days (%)",
    y = "Adults with Obesity (%)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    panel.grid.minor = element_blank()
  )

ggplotly(p_scatter, tooltip = "text")
```

## Trends Over Time: Inactivity and Obesity (2011–2024)

How have physical inactivity and obesity rates changed across states over the past decade? Click **Play** or drag the timeline slider to observe state trajectory over time.

```{r}
#| label: animated-trend
#| echo: false
#| message: false
#| warning: false

trend_data <- readRDS("trend_brfss.rds")

plot_ly(
  data = trend_data,
  x = ~inactivity_pct,
  y = ~obesity_pct,
  text = ~paste0("<b>", state, " (", year, ")</b><br>Inactivity: ", inactivity_pct, "%<br>Obesity: ", obesity_pct, "%"),
  hoverinfo = "text",
  frame = ~year,
  type = "scatter",
  mode = "markers",
  marker = list(size = 11, color = "#2b8cbe", opacity = 0.75, line = list(color = "#08519c", width = 1))
) |>
  layout(
    title = list(text = "U.S. States: Inactivity vs. Obesity Rates (2011–2024)"),
    xaxis = list(title = "Physical Inactivity (%)", range = c(15, 45)),
    yaxis = list(title = "Obesity Rate (%)", range = c(18, 48))
  )
```

## Key Findings

- **Consistent Exercise Advantage:** Across every single age group, adults who reported any physical activity in the past 30 days were roughly twice as likely to rate their health as excellent or very good compared to those reporting no physical activity.
- **Geographic Disparities:** States in the South tend to report higher rates of physical inactivity (often exceeding 28–30%) alongside higher obesity rates, whereas states in the West and New England report lower inactivity rates (often under 20%).
- **Trends Across Age:** While overall self-reported health tends to decline as age increases, the gap between active and inactive individuals remains substantial even among seniors aged 65 and older.
- **Sex Consistency:** Similar patterns hold across both female and male respondents.

