---
title: "Obesity in United States"
---

```{r}
#| label: setup
#| echo: false
#| message: false
#| warning: false

library(tidyverse)
library(plotly)
```

## Geographic Variation Across States

Physical activity levels vary considerably across the United States. The interactive map below shows the percentage of adults reporting no physical activity in the past 30 days, along with state obesity rates.

```{r}
#| label: state-map
#| echo: false
#| message: false
#| warning: false

state_data <- readRDS("state_brfss.rds")

# Configure USA map layout
geo_config <- list(
  scope = "usa",
  projection = list(type = "albers usa"),
  showlakes = TRUE,
  lakecolor = toRGB("white")
)

# Create interactive choropleth map
plot_geo(state_data, locationmode = "USA-states") |>
  add_trace(
    z = ~inactivity_pct,
    text = ~paste0(
      "<b>", state, "</b><br>",
      "Physical Inactivity: ", inactivity_pct, "%<br>",
      "Obesity Rate: ", obesity_pct, "%"
    ),
    hoverinfo = "text",
    locations = ~state_abbr,
    color = ~inactivity_pct,
    colors = "Blues"
  ) |>
  colorbar(title = "Inactivity %") |>
  layout(
    title = list(text = "Physical Inactivity Rate by U.S. State (2023 BRFSS)"),
    geo = geo_config
  )
```

## Inactivity and Obesity Across States

Does higher physical inactivity relate to higher obesity at the state level? The scatter plot below compares the two variables across all 50 states and D.C., with a linear trend line.

```{r}
#| label: scatter-plot
#| echo: false
#| message: false
#| warning: false

p_scatter <- ggplot(
  state_data,
  aes(
    x = inactivity_pct,
    y = obesity_pct,
    text = paste0("<b>", state, "</b><br>Inactivity: ", inactivity_pct, "%<br>Obesity: ", obesity_pct, "%")
  )
) +
  geom_point(color = "#2b8cbe", size = 2.5, alpha = 0.8) +
  geom_smooth(method = "lm", se = FALSE, color = "#e34a33", linetype = "dashed", linewidth = 0.8) +
  labs(
    title = "Physical Inactivity vs. Obesity Rate by State",
    x = "Adults Inactive in Past 30 Days (%)",
    y = "Adults with Obesity (%)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    panel.grid.minor = element_blank()
  )

ggplotly(p_scatter, tooltip = "text")
```

## Trends Over Time: Inactivity and Obesity (2011–2024)

How have physical inactivity and obesity rates changed across states over the past decade? Click **Play** or drag the timeline slider to observe state trajectory over time.

```{r}
#| label: animated-trend
#| echo: false
#| message: false
#| warning: false

trend_data <- readRDS("trend_brfss.rds")

plot_ly(
  data = trend_data,
  x = ~inactivity_pct,
  y = ~obesity_pct,
  text = ~paste0("<b>", state, " (", year, ")</b><br>Inactivity: ", inactivity_pct, "%<br>Obesity: ", obesity_pct, "%"),
  hoverinfo = "text",
  frame = ~year,
  type = "scatter",
  mode = "markers",
  marker = list(size = 11, color = "#2b8cbe", opacity = 0.75, line = list(color = "#08519c", width = 1))
) |>
  layout(
    title = list(text = "U.S. States: Inactivity vs. Obesity Rates (2011–2024)"),
    xaxis = list(title = "Physical Inactivity (%)", range = c(15, 45)),
    yaxis = list(title = "Obesity Rate (%)", range = c(18, 48))
  )
```
